\section{Discussion}
\label{sec:discussion}

\subsection{What the characterization changes}

The characterization positions a state-level investor characteristic in two literatures. In household finance, financial literacy has been studied as a predictor of portfolio quality, market participation, and financial resilience \citep{lusardiMitchell2014, calvetCampbellSodini2007, grinblattKeloharjuLinnainmaa2012}; I position state NFCS literacy as a candidate cross-sectional asset-pricing moderator on a specific anomaly interaction with a non-monotone IO dependence. In the local-return-predictability program built on \citet{korniotisKumar2013}, the contribution adds a literacy moderator alongside the business-cycle channel, distinct but partly overlapping (20--23 percent attenuation under business-cycle controls). Relative positioning to \citet{hillertJacobsMuller2014} and \citet{korniotisKumarPage2020} is in the Introduction.

The mid-IO three-way coefficient on the stable-HQ subsample implies a literacy-modulated return differential of $-7\%$ annualized at a 0.187-standard-deviation swing in the standardized $\mathrm{mom}\cdot\mathrm{IV}\cdot\mathtt{literacy\_z}$ composite. The classical-EIV-corrected coefficient $\hat\beta_{\text{eiv}} = -0.0763$ implies $-16\%$ at the same swing. The relevant economic magnitude is the bracket $-7\%$ to $-16\%$, not a point estimate; the EIV-corrected end is the theoretically principled anchor with $t \in [-2.4, -2.8]$ when $\lambda$ is bootstrapped over the five NFCS waves. The reliability ratio is estimated; a delta-method standard error accounting for wave-bootstrap uncertainty in $\lambda$ would be modestly larger and is not reported. The bracket is one magnitude range on one variation rather than two independent pieces of evidence. It overlaps the $\approx 5$ percent annualized state-business-cycle strategy of \citet{korniotisKumar2013}, with which the literacy moderator is partly correlated.

\subsection{Limitations}

\paragraph{Population scope.} On the corrected PIT full panel the mid-IO coefficient shrinks by a factor of $\approx 26$; the mechanism-strength reading of the collapse is not defendable on current evidence (see Section~\ref{sec:nestedbaseline}). The relocator characteristics (IA~\ref{ia:relocchar}) lean toward the sample-selection reading: relocators are smaller, higher-IV, lower-IO, lower-return, and from slightly lower-literacy states. About one-third of the stable-HQ mid/low gap comes from low-IO attenuation under the relocator drop and about two-thirds from mid-IO growth. I do not run a pre/post-relocation event study on the 1,366 dropped relocators; the stable-HQ restriction is motivated by but not tested against the community-tie argument.

\paragraph{Within-family interpretation.} The within-family interpretation is bounded as in Section~\ref{sec:mech-bounds}: compositional heterogeneity is the design-internal description; the design does not separate it from literacy-correlated local-bias intensity, a mixed-clientele reading, or a no-microfoundation null ($p = 0.94$ persistent, $p = 0.18$ time-varying for the within-mid-IO IV-decile equality test). Brokerage-level data and a within-mid-IO sort on within-firm IO dispersion are the next-paper discriminators not available here. The mid-IO coefficient growth from $-0.0185$ on the full panel to $-0.0321$ on stable-HQ is a joint sample-composition and reweighting effect; I do not decompose it firm-by-firm.

\paragraph{Functional-form fragility of the U-shape.} The continuous-margin quadratic $\hat\gamma_{5} = +0.0151$ ($p = 0.017$) does not survive log-IO ($p = 0.75$), kink-IO ($p = 0.24$), or a flat 20-bin diagnostic (Section~\ref{sec:robust-form}). The trough-location estimate $\mathrm{IO}^{\ast} \approx 0.34$ is correspondingly parametric. The discrete-tercile mid-IO concentration is the headline functional-form-free characterization; the U-shape is its parametric corroboration. The Chow $\chi^{2}(1)\,p = 0.997$ on $\hat\gamma_{5}^{\text{pre}} = \hat\gamma_{5}^{\text{post}}$ is consistent with equality of the curvature across the 2015 break but does not constitute positive structural evidence: the contrast standard error is approximately ninety percent of $\hat\gamma_{5}$ itself (Section~\ref{sec:mech-stability}).

\paragraph{Modern-DiD diagnostics.} I do not run the \citet{callawaySantanna2021}, \citet{sunAbraham2021}, \citet{goodmanBacon2021}, \citet{rambachanRoth2023} diagnostic stack the pre-registered design specified (IA~\ref{ia:did} for the rationale, the underpowering disclosure, and the pre-registered interpretation-rule departure on the realized $+0.0731$ point estimate).

\paragraph{Ownership-interactive state-level confound class.} The within-stock contrast does not discriminate against a state-level moderator that itself operates differently across IO terciles. The canonical instance is state-resident public-pension-fund home-state ownership, where state pensions concentrate holdings in mid-IO firms with mechanical state-level co-movement. A keyword filter of the 13F manager universe identifies 21 distinct state-pension-fund managers concentrated in roughly seven states with active state-pension systems (the modal state-pension state has one or two managers; 44 of 51 states have none in the filter). Even with the 13F home-state-pension-resident control complete, the construction would be quantitatively thin because home-state-pension residence is collinear with the state fixed effect for $\approx 40$ of 51 states. State-employee 401(k) custodians and state-resident-brokerage concentration are the closest extensions of this class; I do not identify a separating test on current data. Closing the class would require a CIK-by-CIK state-pension-resident flag matched against the 13F panel; that data source is not built here.
