\section{Introduction}

This paper characterizes where in the cross-section of institutional ownership (IO) the literacy moderation of the momentum--idiosyncratic-volatility (IV) interaction concentrates. The headline finding is functional-form-free. On the stable-headquarters subsample of CRSP-listed U.S. common stock 2009--2023 (519,090 firm-months, 6,081 firms), stratifying the three-way regression by within-stock IO tercile concentrates the literacy moderation in the mid-IO tercile: $\hat\gamma_{\text{mid}} = -0.0321$ (state-clustered $t = -2.63$; wild-cluster bootstrap $p = 0.064$; family-wise Romano-Wolf $p = 0.077$). The low- and high-IO coefficients are one-third and one-sixth of the mid-tercile magnitude. A continuous-margin quadratic-IO specification corroborates the non-monotone shape with an interior trough at $\mathrm{IO}^{\ast} \approx 0.34$ ($\hat\gamma_{5} = +0.0151$, $p = 0.017$). Literacy dominates the closest education proxy: a state-college-attainment horse-race within mid-IO returns $\hat\gamma_{\text{lit}} = -0.0350$ ($t = -2.97$) and $\hat\gamma_{\text{college}} = +0.00076$ ($t = +0.65$).

The contribution is the discrete-tercile mid-IO concentration on stable-headquarters and the disciplining of the within-family interpretation against the alternatives the design separates and those it does not. The institutional setting is the one in which retail investors overweight geographically proximate firms \citep{covalMoskowitz1999, ivkovicWeisbenner2005} and local investor-base concentration moves prices through a demand channel \citep{hongKubikStein2008}; geographic information asymmetry also shapes institutional behavior \citep{bernileKumarSulaeman2015}. The operative friction is the interaction of limits to arbitrage that scale with idiosyncratic volatility \citep{shleiferVishny1997, pontiff2006} with sophistication-dependent local retail demand modulated by state National Financial Capability Study (NFCS) Big-Five literacy \citep{lusardiMitchell2014, calvetCampbellSodini2009, grinblattKeloharjuLinnainmaa2012}.

The within-family interpretation is held tightly. The mid-IO concentration discriminates against any generic state-level moderator that hits all headquarters-state stocks alike, but it does not select a single within-family microfoundation. Within mid-IO, the headline three-way decomposed by within-month NYSE IV decile fails to reject equality across deciles (joint $\chi^{2}$ $p = 0.94$ persistent; $p = 0.18$ time-varying), and decile 10 (where a canonical limits-to-arbitrage arbitrage-capacity reading predicts the strongest moderation) returns $\hat\gamma \approx 0$. The data therefore rule out the canonical arbitrage-capacity microfoundation, but the IV-decile null cannot distinguish compositional heterogeneity from a no-channel null. I name compositional heterogeneity as the most direct family-internal description of what remains rather than as a positively-identified channel. The quadratic-IO curvature stability across the 2015 break is uninformative under a Chow $\chi^{2}(1)\,p = 0.997$ where the contrast standard error is approximately ninety percent of $\hat\gamma_{5}$ itself; the trough $\mathrm{IO}^{\ast}$ drifts from $0.26$ to $0.40$ as the median IO share rises (Section~\ref{sec:mech-stability}).

Three bounds discipline the result. First, the quadratic five-percent statistical clearance is the central inferential anchor for the parametric corroboration but does not survive log-IO, kink-IO, or a twenty-bin diagnostic; the discrete-tercile contrast is the functional-form-free characterization, and the quadratic provides parametric corroboration only. Second, the mid-IO coefficient collapses 26-fold ($-0.0321 \to -0.00121$) under a point-in-time (PIT) EDGAR 10-K-header literacy reassignment that captures the 14 percent of firm-months whose then-current headquarters state differs from the static Compustat snapshot, and the aggregate three-way collapses approximately 39-fold ($-0.0117 \to -0.0003$), refuting the prior aggregate-three-way prediction; the within-stock retail-versus-institutional difference also sign-flips. The directional within-tercile pattern is preserved on the PIT full panel; the magnitude is not. Third, on the stable-HQ subsample the mid/low ratio is $2.83$ versus a full-panel ratio of $1.09$; relocators are nearly twice as concentrated in the low-IO tercile as elsewhere, and dropping them attenuates low-IO. About one-third of the gap comes from low-IO attenuation and two-thirds from mid-IO growth, so the underlying within-stock contrast on the corrected full panel is smaller than the stable-HQ ratio implies. The corrected sample refutes the prior aggregate prediction; the documented mid-IO concentration on stable-HQ replaces it as the contribution.

The closest prior work is \citet{korniotisKumar2013}, who show that state-level business cycles predict returns of locally headquartered firms. I differ on three margins: the state variable is NFCS literacy on a three-year wave frequency rather than the quarterly business-cycle frequency; the structural form is a three-way moderation of an anomaly interaction; and the discriminating margin is a within-stock IO contrast that the state business-cycle channel does not predict. A business-cycle control block attenuates the aggregate literacy three-way by 20--23 percent, so the two moderators are distinct but partly overlapping; the discrimination claim's scope is generic state-level confounds, not ownership-interactive ones, which remain open. \citet{korniotisKumarPage2020} use brokerage-demographic proxies for state-level sophistication; that paper uses neither NFCS literacy nor the momentum-IV interaction. \citet{hillertJacobsMuller2014} is the closest precedent for a state-level moderator of momentum, with media coverage as the primary moderator and state investor individualism as a cross-sectional moderator of the media-momentum effect, rather than literacy.
