\section{Conclusion}
\label{sec:conclusion}

A large literature documents that less financially literate investors trade systematically worse, and a separate large literature documents that retail demand is geographically concentrated and moves prices. Put those two facts together and the prediction is direct: state-level investor literacy should moderate the return anomalies that depend on retail mispricing. Aggregate tests on U.S. equity panels deliver weak signals. The contribution of this paper is to show that the literacy mechanism is not absent --- it is hidden by aggregation. Where in the cross-section the mechanism operates, and under what conditions on ownership and size, is the question this paper sets out to answer.

The data identify an interaction zone. The literacy channel requires two ingredients to leave a price footprint: retail demand large enough to move prices, and institutional activity sufficient to translate that demand into return differences detectable across literate and illiterate states. Where neither ingredient is present (micro-cap firms held mostly by retail), mispricing accumulates but does not load on state-literacy variation. Where institutions dominate (large firms with high IO), arbitrage absorbs retail demand and neutralizes the local-state channel. The interaction zone --- firms with retail demand at meaningful market size and enough institutional activity to arbitrage at the margin --- is where the literacy channel can be detected, and is where the data find the moderation. In the full stable-HQ panel, the interaction zone falls predominantly in mid-IO. Restricting the panel to larger firms relocates it to low-IO, and the cross-state literacy gradient migrates with it. Firms outside the interaction zone show no gradient at all.

The empirical signature of this story --- migration of the moderation with sample composition --- is the central new piece of evidence the paper introduces. The IO discretization is sample-conditional; the underlying economic population is not. The migration evidence identifies the interaction-zone reading without requiring an instrumental-variables strategy, which is fortunate because the natural candidate instrument (state K-12 personal-finance mandates) fails its first-stage relevance test for documented reasons: low-literacy states adopted mandates in response to low adult literacy, and a cohort-matched implementation also fails. The headline result rests on OLS identification under the conditional-unconfoundedness assumption, and the migration test provides the predicted-signature confirmation that an instrument would otherwise be asked to provide.

Three honest limitations bound the result. The continuous-margin U-shape clears five percent statistical significance on a quadratic specification only; log-IO, kink-IO, and a nonparametric twenty-bin diagnostic all return null curvature. The functional-form-free claim is the discrete-tercile contrast, and the quadratic is corroboration only. The point-in-time literacy reassignment on the full panel collapses the headline coefficient by roughly twenty-six fold, leaving the stable-headquarters subsample as the population for which the characterization holds; this is the binding population-scope limitation. The within-mid-IO IV-decile decomposition fails to reject decile equality, rejecting the textbook reading that arbitrage costs binding at high IV should generate the strongest moderation at the highest IV decile. The microfoundation the paper supports is identified through size and ownership conditioning, not through the IV margin.

The result has implications for how we read state-level investor characteristics in asset pricing. Whether other state-level investor traits --- household participation rates, brokerage density, retail-attention measures --- operate through a similar interaction zone is a natural next question, and the size-restriction migration test developed here provides a portable empirical signature for testing it. Whether the within-family microfoundation is best read as a demand-intensity channel, an attention channel, or a clientele-timing channel is a question the present design treats as observationally similar. The TAQ-based Boehmer-Jones-Zhang retail trade indicator would discriminate among these readings; the data construction for that test is a natural next-paper extension. Pre/post-relocation event studies on the 1{,}366 firms dropped from the stable-HQ sample would discriminate between the mechanism-strength and sample-selection readings of the PIT magnitude collapse. The interaction zone, characterized as it is, suggests where to look next.
