\section{Discussion}
\label{sec:discussion}

\subsection{What the characterization changes}

The characterization positions state NFCS Big-Five financial literacy as a cross-sectional asset-pricing moderator that operates through an interaction-zone microfoundation, identified through the size $\times$ ownership margin rather than the IV margin the canonical limits-to-arbitrage reading predicts. In household finance, financial literacy has been studied as a predictor of portfolio quality, market participation, and financial resilience \citep{lusardiMitchell2014, calvetCampbellSodini2007, grinblattKeloharjuLinnainmaa2012}; this paper positions state-aggregate literacy as a moderator on a specific anomaly interaction whose population scope is precisely characterized. In the local-return-predictability program built on \citet{korniotisKumar2013}, the paper adds a literacy moderator alongside the business-cycle channel; the two state moderators are distinct (the literacy channel is identified through within-stock IO and size conditioning that the business-cycle channel does not predict) but partly overlapping (a business-cycle control block attenuates the aggregate literacy three-way by 20--23 percent). Relative positioning to \citet{hillertJacobsMuller2014} and \citet{korniotisKumarPage2020} is in the Introduction.

The economic magnitude of the result, taking the size-controlled mid-IO coefficient $\hat\gamma_{\text{mid-IO}} = -0.0256$ on the stable-HQ subsample, implies a literacy-modulated annualized return differential of approximately $-5.5\%$ at a 0.187-standard-deviation swing in the standardized $\mathrm{mom}\cdot\mathrm{IV}\cdot\mathtt{literacy\_z}$ composite. The baseline (non-size-controlled) coefficient implies $-7\%$ at the same swing; a classical errors-in-variables correction using the firm-month-weighted reliability ratio $\lambda = 0.420$ (Section~\ref{sec:robustness}) extends the bracket to roughly $-13\%$. The relevant economic magnitude is the bracket $-5.5\%$ to $-13\%$, with the EIV-corrected end as the theoretically principled anchor. The bracket overlaps the $\approx 5\%$ annualized state-business-cycle strategy of \citet{korniotisKumar2013}, with which the literacy moderator is partly correlated.

\subsection{Limitations}

\paragraph{Population scope.} On the corrected PIT full panel the mid-IO coefficient collapses by a factor of $\approx 26$; the mechanism-strength reading of the collapse is not defendable on current evidence. The relocator characteristics (IA~\ref{ia:relocchar}) lean toward a sample-selection reading: relocators are smaller, higher-IV, lower-IO, lower-return firms, and from slightly lower-literacy states. About one-third of the stable-HQ mid/low gap comes from low-IO attenuation under the relocator drop and about two-thirds from mid-IO growth. The interaction-zone characterization in this paper is a statement about firms with stable headquarters across the panel window --- the population for which the local-investor channel is well-defined --- and does not generalize without modification to the full panel under point-in-time literacy reassignment.

\paragraph{Within-family microfoundation.} The interaction-zone characterization is positively identified through the size-restriction migration and the cross-state literacy gradient. It does not separate alternative within-family readings (retail demand intensity, retail attention, retail timing) that make the same predictions for the size $\times$ ownership conditioning but different predictions for retail trade-flow data. The TAQ-based Boehmer-Jones-Zhang retail trade indicator would provide a margin to separate them; the data construction for that test is outside the present panel.

\paragraph{Time variation in the cross-state gradient.} The pooled-panel cross-state literacy gradient (Holm FWER $p = 0.014$) is driven primarily by post-2015 data; the pre-2015 cross-state slope is wrong-signed and concentrated in NFCS Wave 2009. The most likely explanation is NFCS measurement noise in the first administration of the Big-Five module; institutional ownership flows (which would identify a retail-trading regime change) do not correlate with the literacy gradient and the post-2015 IO change is positive (institutional share rose, not fell). The headline cross-state evidence requires the pooled panel for adequate power; the post-2015 sub-sample alone delivers raw-$p = 0.049$ but Holm FWER $p = 0.147$. This is a power issue, not a substantive shift; the within-panel mid-IO coefficient is essentially identical pre- and post-2015.

\paragraph{Instrumental-variables strategy.} The state K-12 personal-finance mandate instrument (Urban-Schmeiser 2015/2020) fails its first-stage relevance test in the cross-state regression of state NFCS literacy on mandate exposure (largest candidate $F = 7.06$, wrong direction). A cohort-matched implementation using NFCS micro-data also fails: with state, wave, and band fixed effects, the cohort-specific first-stage $F = 0.05$. The mandate variable is endogenous to state literacy through a reverse-causality channel (low-literacy states adopted mandates as a remedy). The OLS estimate is the preferred inferential statement. A clean IV strategy would require pre-1990 state economics-education mandate variation \citep{bernheimGarrettMaki2001} or NFCS micro-data restricted to cohorts demonstrably treated under a verifiable-compliance mandate; neither is straightforward.

\paragraph{Ownership-interactive state-level confounds.} The within-stock contrast does not discriminate against a state-level moderator that itself operates differently across IO terciles. The canonical instance is state-resident public-pension-fund home-state ownership, where state pensions concentrate holdings in mid-IO firms with mechanical state-level co-movement. A keyword filter of the 13F manager universe identifies $\approx 21$ distinct state-pension-fund managers concentrated in roughly seven states with active state-pension systems; for the remaining $\approx 44$ states, home-state-pension residence is collinear with the state fixed effect. State-employee 401(k) custodians and state-resident-brokerage concentration are the closest extensions of this confound class. Closing the class requires a CIK-by-CIK state-pension-resident flag matched against the 13F panel; that data source is not built here.

\paragraph{Functional-form fragility of the U-shape.} The continuous-margin quadratic $\hat\gamma_{5} = +0.0151$ ($p = 0.017$) does not survive log-IO ($p = 0.75$), kink-IO ($p = 0.24$), or a flat twenty-bin diagnostic. The trough-location estimate $\mathrm{IO}^{\ast} \approx 0.34$ is correspondingly parametric. The discrete-tercile mid-IO concentration is the headline functional-form-free characterization; the U-shape is its parametric corroboration only.
