\section{Introduction}

\citet{lusardiMitchell2014} and a large literature that follows them document that financially literate households make systematically better portfolio decisions: they diversify more, panic less, time the market less, and pay more attention to fees and volatility-adjusted risk. The variation across U.S. states is substantial. In high-literacy states like Minnesota, Hawaii, and New Hampshire, roughly 28 percent of adults can answer three of five basic literacy questions correctly; in Mississippi, Ohio, and West Virginia, the share is around 21 percent. If the trading decisions of less literate investors leave a footprint in equilibrium prices, that footprint should be visible as a state-level moderator of the return anomalies most plausibly driven by retail behavior --- momentum, idiosyncratic volatility, and their interaction.

The empirical literature on local bias provides the second piece of the setup. \citet{covalMoskowitz1999} and \citet{ivkovicWeisbenner2005} document that retail investors hold portfolios disproportionately weighted toward locally headquartered firms; \citet{hongKubikStein2008} show that this local demand moves prices for the firms in question; \citet{bernileKumarSulaeman2015} extend the result to other dimensions of geographic clientele. The natural prediction follows: where retail investors are less financially literate, they should trade more like the textbook noise trader, and the anomalies their behavior generates should be stronger. State NFCS literacy, applied to firms by headquarters state, gives the moderator; the momentum-IV anomaly interaction, which combines a retail-friendly anomaly with the standard limits-to-arbitrage measure, gives the natural object to study.

Putting these pieces together produces a clean hypothesis: state literacy should moderate the momentum-IV interaction, with the effect weaker in literate states and stronger in illiterate states. In aggregate, however, the data do not deliver. Regressing the three-way interaction on the full panel of CRSP-listed common stock returns a modest negative coefficient with marginal statistical significance, and the magnitude collapses by roughly forty fold once headquarters-state reassignment is corrected for the 14 percent of firm-months whose static Compustat snapshot does not match their then-current 10-K filing. The aggregate test is misleading in two ways. It pools firms with very different ownership structures, and it pools firms whose local-investor base is genuinely defined (firms with stable headquarters across the panel) with firms whose local-investor base is in transition (relocators, for whom neither origin nor destination state correctly characterizes the marginal investor). The aggregate null is therefore not informative about whether a literacy channel exists. It is informative about whether the channel operates uniformly across the cross-section. It does not.

The contribution of this paper is to ask a sharper question. Where in the cross-section does the literacy mechanism become visible? Under what conditions on firm ownership and size does retail mispricing translate into price differences that can be detected across literate and illiterate states? And what does the answer identify about the microfoundation?

The data deliver a sharp pattern. On the stable-headquarters subsample of CRSP-listed common stock from 2009 to 2023, the literacy moderation of the momentum-IV interaction concentrates in firms whose institutional ownership sits in the middle of the within-stock IO distribution --- the mid-IO tercile, with mean ownership share around fifty percent. Firms with predominantly retail ownership (low IO) show essentially no literacy moderation. Firms with predominantly institutional ownership (high IO) also show none. The mid-tercile coefficient survives the addition of firm-size moderation as a control, so the finding is not a size effect dressed in ownership clothing, even though mid-IO firms are predominantly mid-cap. Within mid-IO, the coefficient corresponds to a cross-state literacy gradient that recovers exactly the prediction the hypothesis makes: states with higher NFCS literacy show weaker anomaly slopes, and the across-state correlation between per-state coefficients and state literacy is sharp and statistically clean.

The economic interpretation that the data identify is an interaction zone. The literacy channel requires two ingredients to leave a price footprint: retail demand large enough to move prices, and institutional activity sufficient to translate that demand into return differences detectable across states. Where neither ingredient is present (micro-cap firms held mostly by retail with no institutional counterparty), retail mispricing is genuine but does not load on cross-state literacy variation; it accumulates uniformly without an arbitrage counterpressure. Where institutions dominate (large firms with high IO), institutional arbitrage absorbs whatever retail demand exists and neutralizes the local-state channel; literacy varies, but the marginal investor doesn't. The interaction zone --- the population of firms with mixed ownership at meaningful market size --- is where both ingredients are present, and is where the data find the literacy moderation. In the full stable-HQ panel, the interaction zone falls predominantly in mid-IO. Under sample restrictions to larger firms, it relocates. Specifically: dropping the bottom 50 percent of firms by market capitalization shifts the moderation from mid-IO to low-IO, and the cross-state literacy gradient migrates with it. The same economic population, relocated to a different tercile of the conditional IO distribution.

The migration is the central piece of mechanism evidence the paper introduces. It is the empirical signature an interaction-zone reading predicts and the IO-tercile result alone does not deliver. The high-IO tercile shows no cross-state literacy gradient under any sample restriction, providing a clean predicted-null check the reading passes. The IV-decile decomposition within mid-IO, which would test the canonical limits-to-arbitrage microfoundation under which the effect strengthens at high IV, fails to reject decile equality and returns a near-zero coefficient at the highest IV decile. The textbook limits-to-arbitrage reading is rejected; the interaction-zone reading, identified through size and ownership conditioning rather than through IV, is identified positively by the migration.

Three bounds discipline the result. First, the continuous-margin quadratic-IO five-percent statistical clearance for the U-shape across IO terciles does not survive log, kink, or twenty-bin alternatives; the functional-form-free claim is the discrete-tercile contrast. Second, on the corrected full panel under point-in-time literacy reassignment, the mid-IO coefficient magnitude collapses by roughly twenty-six fold; the characterization is therefore a statement about firms with stable headquarters across the panel window, not about the corrected full panel. Third, an instrumental-variables strategy using state K-12 personal-finance graduation requirements as a source of exogenous variation in adult literacy fails its first-stage relevance test in the right direction --- mandate states are systematically less literate, reflecting that states adopted mandates as a response to low adult literacy. The OLS estimate is the preferred inferential statement, and the IV failure is documented honestly.

The closest prior work is \citet{korniotisKumar2013}, who show that state-level business cycles predict returns of locally headquartered firms. This paper differs on three margins: the state variable is NFCS literacy rather than the business cycle, the object of study is a three-way moderation of an anomaly interaction rather than a level prediction, and the discriminating margin is a within-stock IO contrast that the business-cycle channel does not predict. A business-cycle control block attenuates the literacy three-way by 20--23 percent, so the two state moderators are distinct but partly overlapping. \citet{korniotisKumarPage2020} use brokerage-demographic proxies for state-level sophistication, but neither NFCS literacy nor the momentum-IV interaction. \citet{hillertJacobsMuller2014} is the closest precedent for a state-level moderator of momentum, using media coverage as the primary moderator and state investor individualism as a cross-sectional moderator of the media-momentum effect, rather than literacy.

\paragraph{Roadmap.} Section~\ref{sec:data} describes the data and sample, including the construction of the NFCS state-literacy panel and the stable-headquarters subsample. Section~\ref{sec:identification} states the specification and the identifying assumptions. Section~\ref{sec:results} reports the discrete-tercile mid-IO concentration, the size-controlled extension, the migration, and the cross-state literacy gradient. Section~\ref{sec:mechanism} develops the interaction-zone characterization. Section~\ref{sec:robustness} reports the disciplining battery: the point-in-time correction, the IV-decile decomposition, the pre/post-2015 split, the failed instrumental-variables strategy, and the functional-form grid. Section~\ref{sec:discussion} discusses interpretation and limitations. Section~\ref{sec:conclusion} concludes.
