\section{Robustness}
\label{sec:robustness}

This section reports the disciplining battery. The point-in-time (PIT) headquarters correction that bounds the population scope of the result is the binding check and is reported in full; the remaining legs --- functional-form sensitivity, the pre/post-2015 split, the failed instrumental-variables strategy, the within-retail aggregation-artifact null, the variance-channel diagnostic, the inference procedure, and the wave-anchor and vintage checks --- are summarized here with full detail in the Internet Appendix.

\subsection{The PIT correction and the nested-baseline collapse}
\label{sec:robust_pit}
\label{sec:nestedbaseline}

The seed prior to this paper's revisions predicted a negative aggregate three-way on the full panel and a negative within-stock retail-versus-institutional difference. Table~\ref{tab:nested} reports both objects under the static-snapshot literacy merge and under the point-in-time EDGAR 10-K-header literacy reassignment.

\begin{table}[t]
\centering
\caption{The nested baseline: aggregate and mid-IO three-way under PIT correction}
\label{tab:nested}
\footnotesize
\begin{minipage}{\textwidth}
\centering
\textit{Panel A: Aggregate three-way coefficient}\\[2pt]
\begin{tabular}{lrrrr}
\toprule
Specification & $\hat\gamma$ & State-$t$ & wcb $p$ & $N$ \\
\midrule
Full panel, static-snapshot literacy & $-0.0117$ & $-2.17$ & 0.078 & 623,896 \\
Full panel, PIT literacy reassignment & $-0.0003$ & $-0.86$ & 0.517 & 623,814 \\
Stable-HQ subsample, static literacy  & $-0.0145$ & $-1.94$ & 0.122 & 519,090 \\
\bottomrule
\end{tabular}

\medskip
\textit{Panel B: Mid-IO coefficient under PIT, with and without size controls}\\[2pt]
\begin{tabular}{lrrrr}
\toprule
Specification & $\hat\gamma_{\text{mid-IO}}$ & State-$t$ & Collapse vs.\ baseline & $N$ \\
\midrule
Stable-HQ static, persistent (paper headline) & $-0.0321$ & $-2.63$ & --- & 191,589 \\
Stable-HQ static, size-controlled & $-0.0256$ & $-2.24$ & --- & 191,589 \\
PIT full panel, persistent & $-0.00121$ & $-2.07$ & $\approx 27\times$ smaller & 191,589 \\
PIT full panel, size-controlled & $-0.00100$ & $-1.75$ & $\approx 26\times$ smaller & 191,589 \\
\bottomrule
\end{tabular}
\end{minipage}

\medskip
{\footnotesize\textit{Notes.} Static-snapshot merge pairs each firm-month with the literacy of its Compustat \texttt{state}; PIT reassignment re-merges against the firm's then-current 10-K-header state and re-$z$-scores within month. State-level wild-cluster bootstrap (Webb six-point, $B = 4{,}999$ or $9{,}999$, null imposed). The PIT magnitude collapse is essentially the same with and without size controls (27$\times$ vs.\ 26$\times$).}
\end{table}

Two readings of the collapse compete on the margins this design supports. A \textit{mechanism-strength reading} says the channel operates most strongly on stable-HQ firms, where the local-investor unit of analysis is well-defined, and that the corrected full panel --- including the 14\% of firm-months with relocated headquarters --- dilutes the average. A \textit{sample-selection reading} says the stable-HQ subsample is selected toward firms with stronger anomaly moderation for reasons orthogonal to literacy. The relocator characteristics in IA~\ref{ia:relocchar} (relocators are smaller, higher-IV, lower-IO, lower-return firms) lean toward the sample-selection reading; the mechanism-strength reading is not defendable on current evidence. The interaction-zone characterization is therefore a statement about the stable-HQ population, and generalization to the PIT-corrected full panel is the principal population-scope limitation.

\subsection{Time variation in the cross-state gradient}
\label{sec:robust_prepost}

The within-panel mid-IO three-way coefficient is essentially identical pre- and post-2015 ($\hat\gamma^{\text{pre}} = -0.024$, $\hat\gamma^{\text{post}} = -0.026$; difference $-0.002$, $t = +0.06$): the within-panel direction is stable. What varies across time is the \emph{cross-state} literacy gradient that interprets the result through the literacy mechanism. The pooled gradient is right-signed and sharp ($-0.0623$, state-$t = -5.01$); the post-2015 sub-period is sharper ($-0.0706$, $t = -3.50$); the pre-2015 sub-period is wrong-signed ($+0.0403$, $t = +1.89$) and concentrated in NFCS Wave 2009, the first administration of the Big-Five module (Wave-2009 slope $+0.281$; Waves 2012--2021 directionally negative). The sub-period table and a time-variation figure (rolling-window coefficient negative in every window; pre/post and size-filter panels) are in IA~\ref{ia:prepost} and IA~\ref{ia:gradfigs}. The most parsimonious explanation is NFCS measurement maturation in the 2015+ waves rather than a 2015-era economic regime change: institutional ownership of the panel \emph{rose} post-2015 (mean $\Delta\,\mathtt{io\_share} = +0.083$), ruling out a retail-trading regime shift, and the state-level change in IO is uncorrelated with state literacy ($r = +0.04$).

\subsection{The remaining disciplining battery}
\label{sec:robust_battery}

The remaining legs are reported in full in the Internet Appendix and summarized here.

\textit{Functional form.} The continuous-margin quadratic-IO five-percent clearance ($\hat\gamma_{5} = +0.0151$, $p = 0.017$) does not survive log-IO ($p = 0.75$), kink-IO at the empirical minimum ($p = 0.24$), or a flat twenty-bin diagnostic; the load-bearing characterization is the functional-form-free discrete-tercile contrast of Table~\ref{tab:headline}. The grid is IA~\ref{ia:formgrid}.

\textit{Instrumental variables.} An attempt to instrument state literacy with K-12 personal-finance graduation requirements fails its first-stage relevance test in the right direction: the strongest candidate has $r = -0.355$ (wrong sign) and $F = 7.06$, because low-literacy regions adopted mandates as a remedy, and a cohort-matched implementation with state, wave, and age-band fixed effects returns first-stage $F = 0.05$. 2SLS is directionally consistent with OLS but with standard errors inflated 6--7$\times$; OLS from Equation~\eqref{eq:size_ctrl} is the preferred inferential statement. Full first-stage and 2SLS detail is IA~\ref{ia:ivfirststage}.

\textit{Within-retail null, variance channel, and education.} Within the stable-HQ low-IO tercile, the low-minus-high-literacy difference in the three-way is wrong-signed ($+0.0033$, $t = +2.80$) but decomposes to $\approx 0$ across within-month IV deciles --- an observation-weighting aggregation artifact that provides no positive evidence against the naive sophistication reading (IA~\ref{ia:ivdecile}). The mid-IO concentration is not a variance artifact (mid-IO has \emph{lower} identifying variation than low-IO yet a larger slope), and a state college-attainment horse-race within mid-IO leaves literacy sharp ($-0.0350$, $t = -2.97$) and college null ($+0.00076$, $t = +0.65$): literacy is not a proxy for education.

\textit{Inference, wave-anchor, and vintage.} One-way state clustering on 51 clusters via the wild-cluster bootstrap of \citet{cameronGelbachMiller2008} with Webb six-point weights \citep{webb2014} and the null imposed is the binding procedure throughout.\footnote{The cross-state gradient is interpreted alongside a state-permutation $p$-value ($B = 9{,}999$), with Holm step-down family-wise correction across the three jointly relevant tests; Romano-Wolf step-down is a weakly more conservative upper bound. The two-way state-by-month CGM statistic is reported as a diagnostic in IA~\ref{ia:inference}.} Restricting to the five NFCS wave-anchor years sharpens the mid-IO coefficient ($-0.070$, $t = -2.09$), so the headline is not a forward-fill artifact; fixing each firm's IO tercile to its 2009 assignment attenuates the mid/low ratio from $2.83$ to $2.09$ but preserves the concentration.
